---
title: "Publishing with R"
author: "Daniel Hammarström"
output: html_document
---


## A ggplot2 theme that can be customized



## Creating tables 

Resources for making tables:

[kableExtra](https://cran.r-project.org/web/packages/kableExtra/)

[xtable](https://cran.r-project.org/web/packages/xtable/index.html)

[knitr in a knutshell, Karl Broman](http://kbroman.org/knitr_knutshell/)


### A small example
```{r packages, warning=FALSE, message=FALSE}
library(dplyr) # for data wrangling
library(knitr) # for making tables
library(kableExtra) # for customizing tables
``` 




```{r small example table, warning=FALSE, message=FALSE}
data("PlantGrowth") # load data

# Prepare table of summary statistics
pg_summary <- PlantGrowth %>%
        group_by(group) %>%
        summarise(Mean = round(mean(weight), 2),  # 
                  SD = round(sd(weight), 2))

```

`kable()` is a table generator function in the knitr package. The simple tables generated by `kable()` can be extended using the kableExtra package. The basic table:


```{r table, results='asis'}
options(knitr.table.format = "html") 

kable(pg_summary, # give the prepared data frame to kable
      col.names = c("Group", "Mean", "SD"), # Rename columns of the table
      align = c("l", "c", "r")) # With this option you can align every single column: l, c, r will result in left center and right alignment



```

### Customize with kableExtra
With the knitr rmarkdown output set to HTML we can customize knitr-tables extensively with kableExtra.

#### Changing position, font size and grouping of column headers and row headers

```{r table custom,  results='asis'}

kable(pg_summary, col.names = c("Group", "Mean", "SD"), align = c("l", "c", "r")) %>% # kableExtra can be piped
        
        kable_styling(bootstrap_options = "striped", full_width = FALSE, font_size = 15) %>% 
        
        # kable styling changing apperance, width and font size (more options avaliable) 
        
        add_header_above(c(" " = 1, "Grouping header" = 2)) %>% 
        
        # Add header gioves extra header, e.g. grouping header and group rows produces a row header
        
        group_rows("Grouping of rows", 2, 3) # Header above sepcified rows 
 
     
```

Resource for HTML: [kableExtra HTML](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html)

Resource for PDF: [kableExtra PDF](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_pdf.pdf)







